Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-13 chapters/paper2.tex
Changes 1. Updated text in aim 2 and aim 3 chapters for better flow

This work investigates the structural response of bijels stabilized by anisotropic magnetic particles. Using a Lattice Boltzmann Method 
coupled with Molecular Dynamics, we simulate a binary fluid system where ellipsoidal particles, modeled as magnetic dipoles along their 
symmetry axis, stabilize bijel templates. Magnetic fields applied to bijels made without a magnetic field reveal anisotropic bijel formation, 
with microstructure evolution dependent on particle reordering at the interface, quantified using the six-fold 
Steinhardt order parameter \(\langle Q6 \rangle\).

@@

Bijels have proposed applications as soft matter based material templates, drug delivery vectors and separations systems. The materials 
performance in each application is dependent upon the microstructure of the bijel template. Many of the applications listed above benefit 
from stimuli response, allowing for in-situ microstructure tuning to modify permeability of the material. We want to understand whether 
bijels provide access to reversible structural modifications which would allow stimuli responsive bijels to be used in in-situ applications.

We begin our investigation by characterizing the microstructure changes observed on bijels stabilized with prolate ellipsoids with a 
particle volume fraction of $\phi_p = 0.1$. We apply a constant magnetic field of $\bar{B}_z = 0.2$ onto bijels that were simulated 
with no magnetic field. Once the microstructure has reached its new state, we ramp up the magnetic field to $\bar{B}_z = 0.5$ and 
finally $\bar{B}_z = 1$. We then apply the opposite procedure to reduce the applied field strength, resulting in a magnetic field 
ramp of $\bar{B}_z: 0 \rightarrow 0.2 \rightarrow 0.5 \rightarrow 1 \rightarrow 0.5 \rightarrow 0.2 \rightarrow 0$.We characterize the 
microstructure changes observed through calculating a characteristic length scale, called the average domain size. We calculate the hysteresis 
curve by calculating the domain size at the final timestep of each magnetic field change. We then plot these domain sizes against the magnetic field 
applied to obtain them and split them in the descending and ascending magnetic field directions to demonstrate the differences observed. We plot 
these results in Figure \ref{fig:hysteresis_curve}

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper2/hysteresis_curve.png} 
    \caption{Plot of the hysteresis curve of a bijel stabilized by magnetically responsive prolate particles. We observe that the domain size increases as we increase the applied
    magnetic field strength. However, upon decrease of the field strength, we see that the microstructure does not return to its previous value.} 
    \label{fig:hysteresis_curve} 
\end{figure}

Figure \ref{fig:hysteresis_curve} illustrates the hysteresis behavior of magnetically responsive bijels. We observe a hysteresis response of the 
bijel microstructure, characterized by a non-return of the microstructure to the original average domain size value when reducing the applied 
magnetic field strength. The increase in the average length scale when applying the magnetic field is due to domain coarsening caused by 
reorientation of particles to the magnetic field strength, reducing the interfacial coverage. When decreasing the magnetic field strength, 
we observe that the domain size does not return to the original value meaning that that the microstructure of the bijel is unaffected by 
the reduction in magnetic field. This suggests that the particles at the interface are in a kinetically trapped configuration.

@@

Bijels have proposed applications as soft matter based material
templates, drug delivery vectors and separations systems. The materials
performance in each application is dependent upon the microstructure of
the bijel template. Many of the applications listed above benefit from
stimuli response, allowing for in-situ microstructure tuning to modify
permeability of the material. We want to understand whether bijels
provide access to reversible structural modifications which would allow
stimuli responsive bijels to be used in in-situ applications.
Additionally, in many of the proposed stimuli responsive applications
for bijels, the timescales of response is important to characterize and
would also provide a mechanistic understanding of the reasons why the
microstructure changes occur. We probe these aspects of structural
response of bijels in this section by assessing the response of bijels
stabilized by particles with volume fraction of \(\phi_p = 0.1\)
simulated under no applied field to an applied field in the range
\(\bar{B} = 0, 0.2, 0.5, 1\). We begin our investigation by visualizing
the structure before and after the application of magnetic fields.

@@

In Figure \ref{fig:microstructure_viz-field_on} we show representative
snapshots of the bijel at the initial and final timestep with
\(\bar{B}_z = 0\) and \(\bar{B}_z = 1\). We show that upon application
of the field, particles align to the direction of the field. As the
particles orient to the field, they disrupt the packing of particles at
the interface, causing domain coarsening as the particles rearrange into
new positions. The particles then rejam in place once the interfacial
area and particle cross sectional area match. Domain coarsening observed
in the figure can be explained through the ordering of particles at the
interface, causing a reduction in the available area that the particles
have for stabilization.

To correlate the microstructure evolution to the ordering of the
particle monolayer to the magnetic field, we calculate the nematic order
parameter. \cite{veerman_phase_1992}

@@

In Figure \ref{fig:domain_size-field_on}, we present the time evolution
of the normalized average domain size, where normalization is performed
using the volume-equivalent spherical particle radius \(R = 7.9\). The
results indicate that upon applying a magnetic field to a bijel
initially formed without one, the final average domain size increases by
up to 5.2\% and 2.5\% for oblate- and prolate-particle-stabilized bijels
respectively. Additionally, in the absence of an external field, domain
coarsening occurs for both particle morphologies. However, the results
for oblate particles with a magnetic field is attributed to coalescence
of domains rather than magnetic field driven ordering of the particles.

In the same figure, the time evolution of the nematic order parameter
\(S\) is also plotted to illustrate changes in particle alignment.
Initially, at \(\bar{B}_z = 0\), a gradual increase in \(S\) is
observed, attributed to steric interactions driving particle reordering.
This effect contributes to the domain coarsening seen in the absence of
an applied field, consistent with the findings of Günther et
al.~\cite{gunther_timescales_2014}. In the case of
oblate-particle-stabilized bijels without an external field, domain
coalescence is primarily driven by steric interactions rather than
field-induced rearrangement. Furthermore, the rate and extent of
particle alignment to the field exhibit a dependence on field strength,
as evidenced by the earlier onset of nematic ordering with increasing
\(\bar{B}\). We also observed between the time evolution of the nematic
order parameter and domain size, particularly when comparing results for
\(\bar{B}_z = 0.5, 1\) against those for \(\bar{B}_z = 0.2\). This
suggests that nematic ordering alone may not fully account for the
observed changes in domain size.

Previous studies have demonstrated that microstructural properties,
particularly length scale and tortuosity, play an important role in the
performance of porous materials across various applications, including
battery electrodes and tissue engineering. \cite{zhang_promoting_2019}
\cite{ebner_tortuosity_2014} \cite{hsieh_architected_2021} Specifically,
lower tortuosity and more ordered ion-conducting domains have been shown
to enhance charge/discharge rates in battery systems, while controlled
tortuosity in artificial bone grafts facilitates improved cell growth
and nutrient transport.

Precise characterization of the anisotropic microstructure is essential
for the tailored design of application-specific porous materials. To
quantify the characteristic length scales in each Cartesian direction,
we compute the second-moment-derived structure factor without radial
averaging. \cite{jansen_bijels_2011} We also calculate the tortuosity 
in each cartesian direction. Here, we make a further
refinement simplifying the domain sizes to directions parallel and
orthogonal to the magnetic field. These result in the length scales and
their associated calculation, \(L_{\parallel} = L_{z}\) and
\(L_{\perp} = \frac{L_x + L_y}{2}\). We perform the same simplification
with the directional tortuosities.

Tortuosity has been defined using
the geometric, diffusional and hydraulic tortuosities. Here we define
the diffusional tortuosity, \(\tau = \frac{D}{D_{eff}}\) which
characterizes the ratio between the free diffusion of a tracer and the
effective diffusion of a tracer in the material. Therefore, we plot
the directional second moment derived length scales and tortuosity in
Figure \ref{fig:domain_size_aniso-field_on}.

@@

From Figure \ref{fig:domain_size_aniso-field_on}, we see that domain
size anisotropy is observed upon application of the field for both
particle morphologies. We also see some domain anisotropy with no
applied field, which increases upon application of the magnetic field.
The directional domain sizes, \(L_{\perp}\) and \(L_{\parallel}\)
increase by \(73\%\) and \(7\%\) respectively for oblate particles and
\(10\%\) and \(44\%\) respectively for oblate particles. The axes that
coarsens the most is consistent with past work, explained through the
directions particles reorient in response to the applied magnetic field.
As particles unjam and rejam at the interface due to reorientating to
the magnetic field, they adopt direction specific surface coverages
which manifest as anistropic domains upon jamming. In Figure
\ref{fig:domain_size_aniso-field_on}, we demonstrate that the
application of the magnetic field changes the domain size. The
tortuosity of the bijel is initially around \(\tau \approx 1.5\),
consistent with simulations of the diffusive tortuosities of gyroid
structures using Lattice Boltzmann simulations.
\cite{luo_macroscopic_2020} Upon application of the field,
\(\tau_{\perp}\) slightly decreases while \(\tau_{\parallel}\)
increases. We have seen that an increasing domain size decreases the
tortuosity in our past work. \cite{karthikeyan_formation_2024} This
observation is in line with the domain size changes observed in Figure
\ref{fig:domain_size_aniso-field_on}. The results show that the
tortuosity of the bijel is noticeably affected by the application of
magnetic fields. This data suggests that magnetic fields applied onto
bijels stabilized with ellipsoidal particles undergo domain anisotropy
upon application of magnetic fields. However the particle morphology,
specifically its axis of symmetry, affect the direction that the domain
anisotropy is characterized.

Next, we turn to investigate the time gap between the average domain
size increase and the nematic order parameter. We hypothesized that the
time taken for particle reorientation to the field would control the
degree and rate of domain size change when the bijel begins with
particles with random order. This result suggests that the dynamics of
the system at higher magnetic field strengths are dependent upon other
factors that vary upon application of the magnetic field. To understand
this phenomena, we focus on the dynamics of the particle monolayer.
First, we visualize the particle monolayer of the bijels stabilized with
prolate ellipsoids at the initial and final configuration before and
after application of a magnetic field.

@@

In Figure \ref{fig:particle_viz-field_on}, we see the effect of the
application of the magnetic field on the monolayer. The initial
configuration of particles at the monolayer show that the orientations
are random and that the particles tend to lie flat on the interface.
However, upon application of the magnetic field the particles orient to
the field direction. The effect that Gunther. et al.~observed for bijels
stabilized with anisotropic particles can also be observed in these
snapshots as the particles can be observed to have more local order in
the final state than in the initial state with no applied field. We
observe that for bijels stabilized by prolate ellipsoids, we see the
adoption of more regular arrangements of the colloids on the interface.
We investigate the effect of local particle ordering next. Past work by
Bresme and Faraudo and Davies investigating the reorientation of
particles at interfaces under magnetic fields have shown that a critical
magnetic field exists, above which the particle orients to the field and
below which the particle orientation is a function of the capillary
force and magnetic field derived force. \cite{davies_interface_2014}
\cite{bresme_orientational_2007} Particle adsorption into the interface
with microparticles have also been shown to be irreversible due to the
high adsorption energy. Upon particle reorientation, the microstructure
change can be driven solely through particle unjamming and re-jamming,
or it can be driven through the particles dragging the interface during
this process. The mechanism of unjamming and re-jamming can be
investigated using the average interfacial angle, $\langle \psi \rangle$
and the local ordering of the particles characterized with 
$\langle Q6 \rangle$ Past work has identified that local crystallization 
of colloidal crystals takes place at \(\langle Q6 \rangle = 0.38\). 
\cite{toxvaerd_role_2020}

@@

From Figure \ref{fig:interface_angle-nint-field_on}, we identify that
prolate and oblate particles respond differently at the interface. For
bijels stabilized by oblate ellipsoids, we characterize an increase in
\(\langle \psi \rangle\) from an initial value of \(19 ^{\circ}\),
before it reduces slightly and begins to increase once again. The
initial starting value of \(\langle \psi \rangle\) indicates the
particles are lying close to flat on the interface, before they begin to
tilt out of the interface when the magnetic field is applied. Then,
\(\langle \psi \rangle\) plateaus as the interface begins to move in
response to the tilting particle at the interface. This is caused by the
large adsorption energy of the particles at the interface of the bijel,
caused by high surface tension. Finally, the particles jam in place,
locking the interface at an angle to the particle. We see the rate of
increase in \(\langle \psi \rangle\) at early stages and the time spent
in the plateau increase with a larger applied magnetic field. The final
angle at the interface is also dependent upon the applied field
strength.

When analyzing \(\langle \psi \rangle\) for bijels stabilized by prolate
particles, this behavior can also be observed. \(\langle \psi \rangle\)
begins at \(80 ^{\circ}\). The difference in the initial
\(\langle \psi \rangle\) compared to oblate particles is due to the
direction of the axis of symmetry. In prolate particles, the axis of
symmetry lies parallel with the interface and \(\langle \psi \rangle\)
measures the angle between the interface normal and the orientation of
the axis of symmetry. We see \(\langle \psi \rangle\) decrease slightly,
before plateauing and slowly increasing again. The rate of decrease in
\(\langle \psi \rangle\) is field dependent, although given the small
decrease in \(\langle \psi \rangle\) it is unclear if the final value of
this parameter is field dependent for bijels stabilized with prolate
particles. The changes in \(\langle \psi \rangle\) for bijels stabilized
with prolate particles are small compared to bijels stabilized with
oblate particles. We believe this is due to local particle effects at
the interface. We probe this by identifying the local particle ordering
at the interface characterized by \(\langle Q6 \rangle\).

We see that for oblate particles, \(\langle Q6 \rangle\) sees a
decrease from its initial value. However the dynamics are controlled by
the applied magnetic field strength. We split the phases of dynamics
into three sections, similar to \(\langle \psi \rangle\) with the first
section seeing a decrease in \(\langle Q6 \rangle\) upon application of
the magnetic field, followed by a plateau and finally a gradual increase
before jamming takes place. These results indicate that the ordering of
the particles is reduced by the applied magnetic field, before a
restoration of some of the pre-existing order. The final values of
\(\langle Q6 \rangle\) are magnetic field dependent, with a stronger
magnetic field having a lower final \(\langle Q6 \rangle\) value
indicating that either the ordering of particles influence the angle at
the interface or vice versa. We observe that for bijels stabilized by
oblate particles, the interface ordering of the particles increases as a
function of the applied field strength. We also observe from plotting
\(\langle Q6 \rangle\) that application of a magnetic field creates a
local crystallization of the colloid monolayer characterized as
\(\langle Q6 \rangle > 0.38\). \cite{toxvaerd_role_2020}

Figure \ref{fig:domain_size-field_on}, we observed a gap between the
growth of the nematic order parameter and an increase in the domain
size. To probe why this occurs, we recast \(\langle \psi \rangle\),
\(\langle Q6 \rangle\) and \(S\) in terms of the physical phenomena
they represent. The nematic order parameter represents the ordering of
particles to an average direction, which in the case of the application
of a magnetic field, is the field direction. Therefore, a timescale
reliant on the nematic order parameter would represent a timescale of
response rate limited to the response to the magnetic field, which
indicates a particle-field orientation dependence. We name this
timescale \(\tau_S\). When applying a field to a bijel system with a
randomly oriented particle monolayer, the most significant transition
point is the isotropic to nematic transition or when \(S \geq 0.3\). We
define \(\tau_S\) for this section of the analysis by identifying when
\(S \geq 0.3\).

We can also do the same for \(\langle \psi \rangle\) and
\(\langle Q6 \rangle\). A timescale derived from
\(\langle \psi \rangle\) named \(\tau_{\langle \psi \rangle}\)
represents the capillary timescale between the interface and the
particle. Given the shape of the plot, we assign the
\(\tau_{\langle \psi \rangle}\) to be the time when the first minimum of
\(\langle \psi \rangle\) is observed for prolate particles and the first
maximum for oblate particles. A timescale derived from
\(\langle Q6 \rangle\), named \(\tau_{\langle Q6 \rangle}\) is also
calculated which represents a timescale of local particle rearrangement.
We define \(\tau_{\langle Q6 \rangle}\) as the time when
\(\langle Q6 \rangle \geq 0.38\) for prolate particles and the global
minimum of \(\tau_{\langle Q6 \rangle}\) for the oblate particles. We
omit \(\bar{B} = 0\) as previous results have identified the domain size
changes observed there. We then plot the average domain size \(L\)
against time rescaled with these three timescales in Figure
\ref{fig:domain_size-field_on-scaled}.

@@

In Figure \ref{fig:domain_size-field_on-scaled}, the left plot plots the
average domain size against time rescaled using \(\tau_S\). We see that
the domain sizes do not collapse onto a single line, which is expected
if we want to explain the dynamics using the nematic order parameter.
This confirms the observation made above that the dynamics do not only
depend on the reorientation of particles to the field. The middle plot
demonstrates the average domain size against time rescaled using
\(\tau_{\langle \psi \rangle}\). We see that there is good agreement
between all simulations for prolate particles, although the point when
\(\frac{t}{\tau_{\langle \psi \rangle}} = 1\) lies when the domain size
transition has started. However for oblate particles we see that this
timescale does not describe the dynamics well.

The right plot demonstrates the domain size change after rescaling time
with \(\tau_{\langle Q6 \rangle}\) where we see good collapse between
all plots while the domain size transition also begins at
\(\frac{t}{\tau_{\langle Q6 \rangle}} = 1\). The latter point
demonstrates that the dynamics for the structural response of randomly
oriented particles are controlled by the local particle rearrangement at
the interface rather than the orientation of the particle to the
magnetic field.We see that before \(\tau_{\langle Q6 \rangle} = 1\),
\(\bar{B} = 0.2\) for the oblate particles does not collapse onto the
other plots, indicating that the dynamics before that point are
controlled by another phenomena.

>>>>

This work investigates the structural response of bijels stabilized by anisotropic magnetic particles. Using a Lattice Boltzmann Method 
coupled with Molecular Dynamics, we simulate a binary fluid system where ellipsoidal particles, modeled as magnetic dipoles along their 
symmetry axis, stabilize bijel templates. Magnetic fields applied to bijels made without a magnetic field reveal anisotropic bijel formation, 
with microstructure evolution dependent on particle reordering at the interface, quantified using the six-fold 
Steinhardt order parameter $\langle Q6 \rangle$.

@@

Bijels have proposed applications as soft matter based material templates, drug delivery vectors and separations systems. The materials 
performance in each application is dependent upon the microstructure of the bijel template. Many of the applications listed above benefit 
from stimuli response, allowing for in-situ microstructure tuning to offer reversible modifications of material permeability. In this section, we analyzes
whether stimuli responsive bijels provide access to reversible structural modifications.

We begin our investigation by characterizing the microstructure changes observed on bijels stabilized with prolate ellipsoids with a 
particle volume fraction of $\phi_p = 0.1$. We only focus on one system here as the mechanism for stimuli response shown in Aim 1 is identical
for both systems. Therefore we expect that the qualitative trends between prolate and oblate particles will be matched. We select prolate particles
as they are more readily synthesized in experiments.

We apply a constant magnetic field of $\bar{B}_z = 0.2$ onto bijels that were simulated 
with no magnetic field. Once the microstructure has reached its new state, we ramp up the magnetic field to $\bar{B}_z = 0.5$ and 
finally $\bar{B}_z = 1$. We then apply the opposite procedure to reduce the applied field strength, resulting in a magnetic field 
ramp of $\bar{B}_z: 0 \rightarrow 0.2 \rightarrow 0.5 \rightarrow 1 \rightarrow 0.5 \rightarrow 0.2 \rightarrow 0$.We characterize the 
microstructure changes observed through calculating a characteristic length scale, called the average domain size. We calculate the hysteresis 
curve by calculating the domain size at the final timestep of each magnetic field change. We then plot these domain sizes against the magnetic field 
applied to obtain them and split them in the descending and ascending magnetic field directions to demonstrate the differences observed. We plot 
these results in Figure \ref{fig:hysteresis_curve}

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper2/hysteresis_curve.png} 
    \caption{Plot of the hysteresis curve of a bijel stabilized by magnetically responsive prolate particles. We observe that the domain size increases as we 
    increase the applied magnetic field strength. However, upon decrease of the field strength, we see that the microstructure does not return to its previous value,
    demonstrating microstructure hysteresis.} 
    \label{fig:hysteresis_curve} 
\end{figure}

Figure \ref{fig:hysteresis_curve} we observe a hysteresis response of the bijel microstructure, characterized by a difference in the average domain size
when increasing and decreasing the magnetic field. The increase in the average length scale when applying the magnetic field is due to domain coarsening caused by 
reorientation of particles to the magnetic field strength, reducing the interfacial coverage. When decreasing the magnetic field strength, 
we observe that the domain size does not return to the original value meaning that that the microstructure of the bijel is unaffected by 
the reduction in applied magnetic field.

This behavior is also observed in past work investigating the shape deformation of particle stabilized emulsions under 
electric fields. \cite{cui_stabilizing_2013} Cui et al. demonstrated that the particles unjammed and rejammed into a new kinetically arrested state resulting in
the droplet being elongated in the direction of the applied field. This suggests that the particles at the interface of magnetically responsive bijels are also 
in a kinetically trapped configuration. We observe in Figure \ref{fig:hysteresis_curve} that the response of the bijel is dependent upon the field strength applied.
In the following sections, we investigate the impact of magnetic field strength and the initial ordering of the particle monolayer on the observed stimuli response.

@@

From the results in Figure \ref{fig:hysteresis_curve}, we show that the domain size change is dependent upon the applied magnetic field. As porous materials templates,
the microstructure of the material is of interest as it controls the performance of the material. Additionally, in many of the proposed stimuli responsive applications
for bijels, the timescales of response is important to characterize and would also provide a mechanistic understanding of the reasons why the microstructure changes 
occur. We probe these aspects of structural response of bijels in this section by assessing the response of bijels stabilized by particles with volume fraction of 
$\phi_p = 0.1$ simulated under no applied field to an applied field in the range $\bar{B} = 0, 0.2, 0.5, 1$. We begin our investigation by visualizing the structure before 
and after the application of magnetic fields.

@@

In Figure \ref{fig:microstructure_viz-field_on} we show representative snapshots of the bijel at the initial and final timestep with $\bar{B}_z = 0$ and $\bar{B}_z = 1$. 
We show that upon application of the field, particles align to the direction of the field. As the particles orient to the field, they disrupt the packing of particles at
the interface. This caused domain coarsening as the particles rearrange into new positions before rejamming in place once the interfacial
area and particle cross sectional area match. Domain coarsening observed in the figure can be explained through the ordering of particles at the interface, causing a 
reduction in the available area that the particles have for stabilization. To characterize the length scale of the microstructure upon application of the magnetic field,
we calculate the average domain size. To correlate the microstructure evolution to the ordering of the particle monolayer to the magnetic field, we calculate the 
nematic order parameter. \cite{veerman_phase_1992}

@@

In Figure \ref{fig:domain_size-field_on}, we present the time evolution of the normalized average domain size, where normalization is performed
using the volume-equivalent spherical particle radius $R = 7.9$. The results indicate that upon applying a magnetic field to a bijel
initially formed without one, the final average domain size increases by up to 5.2\% and 2.5\% for oblate- and prolate-particle-stabilized bijels
respectively. The domain coarsening is initially slow, before accelerating rapidly and finally reaching a maximum where it plateaus. 
Additionally, in the absence of an external field, domain coarsening occurs for both particle morphologies. However, the results
for oblate particles with a magnetic field is attributed to coalescence of domains rather than magnetic field driven ordering of the particles.

In the same figure, the time evolution of the nematic order parameter $S$ is also plotted to illustrate changes in particle alignment.
Initially, at $\bar{B}_z = 0$, a gradual increase in $S$ is observed, attributed to steric interactions driving particle reordering.
This effect contributes to the domain coarsening seen in the absence of an applied field, consistent with the findings of Günther et
al.~\cite{gunther_timescales_2014}. In the case of oblate-particle-stabilized bijels without an external field, domain
coalescence is primarily driven by steric interactions rather than field-induced rearrangement. Furthermore, the rate and extent of
particle alignment to the field exhibit a dependence on field strength, as evidenced by the earlier onset of nematic ordering with increasing
$\bar{B}$. We also observed between the time evolution of the nematic order parameter and domain size, particularly when comparing results for
$\bar{B}_z = 0.5, 1$ against those for $\bar{B}_z = 0.2$. This suggests that nematic ordering alone may not fully account for the
observed changes in domain size.

Previous studies have demonstrated that microstructural properties such as the length scale and tortuosity, play an important role in the
performance of porous materials across various applications, including battery electrodes and tissue engineering. \cite{zhang_promoting_2019}
\cite{ebner_tortuosity_2014} \cite{hsieh_architected_2021} Specifically, lower tortuosity and more ordered ion-conducting domains have been shown
to enhance charge/discharge rates in battery systems, while controlled tortuosity in artificial bone grafts facilitates improved cell growth
and nutrient transport.

Precise characterization of the anisotropic microstructure is essential for the tailored design of application-specific porous materials. To
quantify the characteristic length scales in each Cartesian direction, we compute the second-moment-derived structure factor. \cite{jansen_bijels_2011} 
We also calculate the tortuosity  in each cartesian direction. Here, we make a further
refinement simplifying the domain sizes to directions parallel and orthogonal to the magnetic field. These result in the length scales and
their associated calculation, $L_{\parallel} = L_{z}$ and $L_{\perp} = \frac{L_x + L_y}{2}$. We perform the same simplification
with the directional tortuosities.

Tortuosity has been defined using the geometric, diffusional and hydraulic tortuosities. Here we define
the diffusional tortuosity, $\tau = \frac{D}{D_{eff}}$ which characterizes the ratio between the free diffusion of a tracer and the
effective diffusion of a tracer in the material. Therefore, we plot the directional second moment derived length scales and tortuosity in
Figure \ref{fig:domain_size_aniso-field_on}.

@@

From Figure \ref{fig:domain_size_aniso-field_on}, we see that domain size anisotropy is observed upon application of the field for both
particle morphologies. We also see some domain anisotropy with no applied field, which increases upon application of the magnetic field.
The directional domain sizes, $L_{\perp}$ and $L_{\parallel}$ increase by 73\% and 7\%\ respectively for oblate particles and
10\% and 44\% respectively for oblate particles. The axes that coarsens the most is consistent with past work, explained through the
directions particles reorient in response to the applied magnetic field.

As particles unjam and rejam at the interface due to reorientating to the magnetic field, they adopt direction specific surface coverages
which manifest as anistropic domains upon jamming. In Figure \ref{fig:domain_size_aniso-field_on}, we demonstrate that the
application of the magnetic field changes the domain size. The tortuosity of the bijel is initially around $\tau \approx 1.5$,
consistent with simulations of the diffusive tortuosities of gyroid structures using Lattice Boltzmann simulations.
\cite{luo_macroscopic_2020} Upon application of the field, $\tau_{\perp}$ slightly decreases while $\tau_{\parallel}$
increases. We have seen that an increasing domain size decreases the tortuosity in our past work. \cite{karthikeyan_formation_2024} This
observation is in line with the domain size changes observed in Figure \ref{fig:domain_size_aniso-field_on}. The results show that the
tortuosity of the bijel is noticeably affected by the application of magnetic fields. This data suggests that magnetic fields applied onto
bijels stabilized with ellipsoidal particles undergo domain anisotropy upon application of magnetic fields. However the particle morphology,
specifically its axis of symmetry, affect the direction that the domain anisotropy is characterized.

Next, we return to investigate the time gap between the average domain size increase and the nematic order parameter. This result suggests that the dynamics 
of the system at higher magnetic field strengths are dependent upon other factors that vary upon application of the magnetic field. In this system, these factors
would arise from capillary forces with the interface or steric interactions with other particles at the interface, slowing down the dynamics of the system.
To understand this phenomena, we focus on the dynamics of the particle monolayer. First, we visualize the particle monolayer of the bijels stabilized with prolate 
ellipsoids at the initial and final configuration before and after application of a magnetic field.

@@

In Figure \ref{fig:particle_viz-field_on}, we see the initial configuration of particles at the monolayer show that the orientations are random and that the 
particles tend to lie flat on the interface. However, upon application of the magnetic field the particles orient to the field direction. With no applied field,
we see that the orientations and positions of the particles vary slightly for both ellipsoidal particle morphologies. This effect is due steric effects between particles,
as identified in Gunther et al. \cite{gunther_timescales_2014} We observe that for bijels stabilized by prolate ellipsoids, we see the adoption of orientational order
to the magnetic field and qualitatively different particle arrangements on the interface. These indicate that the interfacial capillary forces or steric interactions 
with other particles differ upon application of the magnetic field. The particles would prefer to lie flat on the interface to minimize the interfacial energy of 
adsorption onto the interface. Changes in the local arrangement arise from the unjamming and rejamming process upon application of the magnetic field. These differences 
in the orientational and interfacial ordering can be characterized to understand how each respective interaction force changes as a function of time.

Bresme and Faraudo and Davies investigating the reorientation of particles at interfaces under magnetic fields have shown that a critical magnetic field exists, 
above which the particle orients to the field and below which the particle orientation is a function of the capillary force and magnetic field derived force. 
\cite{davies_interface_2014} \cite{bresme_orientational_2007} Particle adsorption into the interface with microparticles have also been shown to be irreversible due 
to the high adsorption energy. Upon particle reorientation, the microstructure change can be driven solely through particle unjamming and re-jamming, or it can be 
driven through the particles dragging the interface during this process. The mechanism of unjamming and re-jamming can be investigated using the average interfacial 
angle, $\langle \psi \rangle$ and the local ordering of the particles characterized with  $\langle Q6 \rangle$ Past work has identified that local crystallization 
of colloidal crystals takes place at \(\langle Q6 \rangle = 0.38\). \cite{toxvaerd_role_2020}

@@

From Figure \ref{fig:interface_angle-nint-field_on}, we identify that prolate and oblate particles respond differently at the interface. For
bijels stabilized by oblate ellipsoids, we characterize an increase in $\langle \psi \rangle$ from an initial value of \(19 ^{\circ}\),
before it reduces slightly and begins to increase once again. The initial starting value of $\langle \psi \rangle$ indicates the
particles are lying close to flat on the interface, before they begin to tilt out of the interface when the magnetic field is applied. Then,
$\langle \psi \rangle$ plateaus as the interface begins to move in response to the tilting particle at the interface. This is caused by the
large adsorption energy of the particles at the interface of the bijel, caused by high surface tension. Finally, the particles jam in place,
locking the interface at an angle to the particle. We see the rate of increase in $\langle \psi \rangle$ at early stages and the time spent
in the plateau increase with a larger applied magnetic field. The final angle at the interface is also dependent upon the applied field
strength.

When analyzing $\langle \psi \rangle$ for bijels stabilized by prolate particles, this behavior can also be observed. $\langle \psi \rangle$
begins at $10 ^{\circ}$.We see $\langle \psi \rangle$ decrease slightly, before plateauing and slowly increasing again. The rate of decrease in 
$\langle \psi \rangle$ is field dependent, although given the small decrease in $\langle \psi \rangle$ it is unclear if the final value of this parameter 
is field dependent for bijels stabilized with prolate particles. The changes in $\langle \psi \rangle$ for bijels stabilized with prolate particles are 
small compared to bijels stabilized with oblate particles. We believe this is due to local particle effects at the interface. We probe this by identifying 
the local particle ordering at the interface characterized by $\langle Q6 \rangle$.

We see that for oblate particles, $\langle Q6 \rangle$ sees a decrease from its initial value. However the dynamics are controlled by
the applied magnetic field strength. We split the phases of dynamics into three sections, similar to $\langle \psi \rangle$. In the first
section, we see a decrease in $\langle Q6 \rangle$ upon application of the magnetic field, followed by a plateau and finally a gradual increase
before jamming takes place. These results indicate that the ordering of the particles is reduced by the applied magnetic field, before a
restoration of some of the pre-existing order. The final values of $\langle Q6 \rangle$ are magnetic field dependent, with a stronger
magnetic field having a lower final $\langle Q6 \rangle$ value indicating that either the ordering of particles influence the angle at
the interface or vice versa. We observe that for bijels stabilized by oblate particles, the interface ordering of the particles increases as a
function of the applied field strength. We also observe from plotting $\langle Q6 \rangle$ that application of a magnetic field creates a
local crystallization of the colloid monolayer characterized as $\langle Q6 \rangle > 0.38$. \cite{toxvaerd_role_2020}

Next, we want to identify which of these forces represent the timescale that controls the dynamics of the system. We can calculate a characteristic
timescale for each factor from the calculated particle properties.
The nematic order parameter $S$ represents the ordering of
particles to an average direction, which in the case of the application
of a magnetic field, is the field direction. Therefore, a timescale
reliant on the nematic order parameter would represent a timescale of
response rate limited to the response to the magnetic field, which
indicates a particle-field orientation dependence. We name this
timescale $\tau_S$. When applying a field to a bijel system with a
randomly oriented particle monolayer, the most significant transition
point is the isotropic to nematic transition or when $S \geq 0.3$. We
define $\tau_S$ as the the isotropic to nematic transition point.

We can also do the same for $\langle \psi \rangle$ and
$\langle Q6 \rangle$. A timescale derived from
$\langle \psi \rangle$ named $\tau_{\langle \psi \rangle}$
represents the capillary timescale between the interface and the
particle. Given the shape of the plot, we assign the
$\tau_{\langle \psi \rangle}$ to be the time when the first minimum of
$\langle \psi \rangle$ is observed for prolate particles and the first
maximum for oblate particles. A timescale derived from
$\langle Q6 \rangle$, named $\tau_{\langle Q6 \rangle}$ is also
calculated which represents a timescale of local particle rearrangement.
We define $\tau_{\langle Q6 \rangle}$ as the time when
\(\langle Q6 \rangle \geq 0.38\) for prolate particles and the global
minimum of $\tau_{\langle Q6 \rangle}$ for the oblate particles. We
omit \(\bar{B} = 0\) as previous results have identified the domain size
changes observed there. We then plot the average domain size \(L\)
against time rescaled with these three timescales in Figure
\ref{fig:domain_size-field_on-scaled}.

@@

In Figure \ref{fig:domain_size-field_on-scaled}, the left plot plots the
average domain size against time rescaled using $\tau_S$. We see that
the domain sizes do not collapse onto a single line, which is expected
if we want to explain the dynamics using the nematic order parameter.
This confirms the observation made above that the dynamics do not only
depend on the reorientation of particles to the field. The middle plot
demonstrates the average domain size against time rescaled using
$\tau_{\langle \psi \rangle}$. We see that there is good agreement
between all simulations for prolate particles, although the point when
$\frac{t}{\tau_{\langle \psi \rangle}} = 1$ lies when the domain size
transition has started. However for oblate particles we see that this
timescale does not describe the dynamics well.

The right plot demonstrates the domain size change after rescaling time
with $\tau_{\langle Q6 \rangle}$ where we see good collapse between
all plots while the domain size transition also begins at
$\frac{t}{\tau_{\langle Q6 \rangle}} = 1$. The latter point
demonstrates that the dynamics for the structural response of randomly
oriented particles are controlled by the local particle rearrangement at
the interface rather than the orientation of the particle to the
magnetic field. As the dynamics of both particle morphologies are well described
using the particle steric effects timescale, denoted by $\tau_{\langle Q6 \rangle}$,
we conclude that the dynamics of bijels stabilized by ellipsoidal particles are controlled by
interfacial particle rearrangement at the interface.

From the evidence in Figure \ref{fig:domain_size-field_on-scaled}, two
types of interfacial behaviors are characterized in addition to that
identified by Gunther et al.~\cite{gunther_timescales_2014} The first
type is what is observed for $\bar{B} = 0.2$, where the interfacial
coverage and average orientation of the particle to the field and to the
interface change in lock step. This causes the interface to reorient at
the same rate as the particles reorienting to the magnetic field. The
second is observed for $\bar{B} = 0.5, 1$, where there is a time lag
between particle ordering, and the intiation of domain coarsening. In
the latter scenario, at early times the particles are jolted out of
position before the large adsorption energy of the particles into the
interface causes the interface to move to the particle position, causing
domain coarsening as the interfacial area reduces.
